Tool-driven Revolutions in Archaeological Science

### 2018-12-13 analysis/paper/jcaa-paper.Rmd
addesd jcaa.csl
but does not seem to be correct. in their description they say they want full first names, not just first letter. and dots seem to be missing.
pity. i don't know anything about csl...

---
title: "Tool-driven Revolutions in Archaeological Science"
author:
  - Sophie C. Schmidt
  - Ben Marwick
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
abstract: "There is an argument in philosophy of science that revolutions in science are either ideas-driven or tool-driven. We explore this debate in light of recent efforts by many scientific disciplines to embrace methods to improve the reproducibility of their research. One of the most profound changes driven by this concern for reproducibility and transparency is from analysing data using tools dependent on pointing-and-clicking with a mouse in commercial software, to a new set of tools based on writing scripts in free and open source programming languages. We review bibliometric evidence for this change in the sciences generally and in archaeology in particular to test if the adoption of these new tools may be considered a revolution or some other process. We evaluate how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches."
---

@@

# tally by year
cran_cites_all_areas_top_journals_by_year <- 
  cran_cites_all_areas_all_years %>% 
  filter(SO %in% top_journals_for_cran_cites$SO[1:10]) %>% 
  group_by(SO, PY) %>% 
  tally(sort = TRUE) %>% 
  mutate(journalname = str_remove_all(SO, " |-"),
         PY = as.integer(PY)) %>% 
  left_join(top_journals_for_cran_cites_articles_per_year,
            by = c('journalname' = 'X4', 
                   'PY' = "year")) %>% 
  mutate(prop = n ) %>% # dropped / value to not have percentages
  filter(SO != 'JOURNAL OF STATISTICAL SOFTWARE') %>% 
  ungroup()

@@

```{r ecology-articles-citing-r-plot, fig.cap=str_glue('Articles per year citing R in top Ecology journals ({prettyNum(articles_shown_in_plot,",")} articles out of {prettyNum(how_many_articles_in_top_journals_for_cran_cites,",")}). Data from Web of Science for 2007-2017')}
min_y <- 2007
max_y <- 2050
all_sciences_citing_R_over_time <- 
ggplot(cran_cites_all_areas_top_journals_by_year,
       aes(PY,
           prop,
           colour = SO)) +
  geom_line(size = 2) +
  geom_text_repel(
    data = subset(cran_cites_all_areas_top_journals_by_year, 
                  PY == max(PY)),
    aes(label = str_wrap(SO_n, 40)),
    size = 4,
    nudge_x = 0,
    hjust = 0,
    segment.color = NA,
    direction = "y"
  ) +
   scale_x_continuous(breaks = min_y:max_y,
                     labels = c(min_y:2017, rep("", length(2018:max_y))),
                     limits = c(min_y, max_y)) +
#  scale_y_continuous(labels = scales::percent) +
  xlab("Publication year") +
  ylab("Number of Articles in journal per year") +
  theme_minimal() +
  theme(legend.position="none",
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

@@

# tally by year
archy_cran_cites_all_areas_top_journals_by_year <- 
  archy_cran_all_years %>% 
  filter(SO %in% top_journals_for_archy$SO) %>% 
  group_by(SO, PY) %>% 
  tally(sort = TRUE)  %>% 
  left_join(top_journals_for_archy_per_year,
            by = c('SO', 'PY')) %>% 
   mutate(prop = n.x) %>% # dropped /n.y to not have percentage
  ungroup()

@@

# filter so not to have the "just one article in R"-journals clutter up the graph

archy_cran_cites_all_areas_top_journals_by_year <- 
archy_cran_cites_all_areas_top_journals_by_year %>%
    filter(total_n_articles > 1)

min_y <- 2007
max_y <- 2050

all_archaeology_citing_R_over_time <- 
ggplot(archy_cran_cites_all_areas_top_journals_by_year,
       aes(PY,
           prop,
           colour = SO)) +
  geom_line(size = 2) +
  geom_point() +
  geom_text_repel(
    data = archy_cran_cites_all_areas_top_journals_by_year %>% 
      group_by(SO) %>% 
      slice(which.max(PY)) , 
    aes(label = SO_n),
    direction = "y",
    size = 5,
    nudge_x = 1,
    hjust = 0) +
  theme_minimal() +
  theme(legend.position="none",
        axis.text.x = element_text(angle = 90, 
                                   hjust = 1, 
                                   vjust = 0.5)) +
  scale_x_continuous(breaks = min_y:max_y,
                     labels = c(min_y:2017, 
                                rep("",
                                    length(2018:max_y))),
                     limits = c(min_y, max_y)) +
  #scale_y_continuous(labels = scales::percent,
   #                  limits = c(0, 0.2)) +
  xlab("Publication year") +
  ylab("Number of Articles in journal per year")

@@

jas_cites_r <- 
archy_cran_cites_all_areas_top_journals_by_year %>% 
  filter(SO == "JOURNAL OF ARCHAEOLOGICAL SCIENCE") %>% 
  ggplot(aes(PY,
           prop)) +
  geom_point(size = 4) +
  geom_smooth(method = "lm",
              se = FALSE) +
  stat_poly_eq(aes(label =  paste(..eq.label.., 
                                  ..adj.rr.label..,
                                  sep = "~~~~")),
               formula = y~x, 
               parse = TRUE,
               size = 3,
               label.y.npc = 0.7) +
  stat_fit_glance(label.y.npc = 0.7,
                  size = 3,
                  method = "lm", 
                  method.args = list(formula = y ~ x),
                  geom = "text",
                  aes(label = paste("p-value: ",
                                    signif(..p.value.., 
                                           digits = 4)))) +
  xlab("year") +
  ylab("Number of articles citing R") +
  ggtitle(expression(paste("Articles citing R in ", 
                           italic("Journal of Archaeological Science")))) +
  theme_bw(base_size = 6)
```

@@

# combine main plot and subplot
all_archaeology_citing_R_over_time_and_subplot <- 
all_archaeology_citing_R_over_time  +
  annotation_custom(grob = ggplotGrob(jas_cites_r), 
                          xmin = 2030, xmax = 2045, 
                          ymin = 12, ymax = 17) # changed position because of numbers now instead of percentage

>>>>

---
title: "Tool-driven Revolutions in Archaeological Science"
author:
  - Sophie C. Schmidt
  - Ben Marwick
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: journal-of-computer-applications-in-archaeology.csl
abstract: "There is an argument in philosophy of science that revolutions in science are either ideas-driven or tool-driven. We explore this debate in light of recent efforts by many scientific disciplines to embrace methods to improve the reproducibility of their research. One of the most profound changes driven by this concern for reproducibility and transparency is from analysing data using tools dependent on pointing-and-clicking with a mouse in commercial software, to a new set of tools based on writing scripts in free and open source programming languages. We review bibliometric evidence for this change in the sciences generally and in archaeology in particular to test if the adoption of these new tools may be considered a revolution or some other process. We evaluate how computational approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches."
---

@@

# tally by year
cran_cites_all_areas_top_journals_by_year <- 
  cran_cites_all_areas_all_years %>% 
  filter(SO %in% top_journals_for_cran_cites$SO[1:10]) %>% 
  group_by(SO, PY) %>% 
  tally(sort = TRUE) %>% 
  mutate(journalname = str_remove_all(SO, " |-"),
         PY = as.integer(PY)) %>% 
  left_join(top_journals_for_cran_cites_articles_per_year,
            by = c('journalname' = 'X4', 
                   'PY' = "year")) %>% 
  mutate(prop = n / value ) %>% 
  filter(SO != 'JOURNAL OF STATISTICAL SOFTWARE') %>% 
  ungroup()

@@

```{r ecology-articles-citing-r-plot, fig.cap=str_glue('Number of articles per year citing R in top Ecology journals ({prettyNum(articles_shown_in_plot,",")} articles out of {prettyNum(how_many_articles_in_top_journals_for_cran_cites,",")}). Data from Web of Science for 2007-2017')}
min_y <- 2007
max_y <- 2025
all_sciences_citing_R_over_time <- 
ggplot(cran_cites_all_areas_top_journals_by_year,
       aes(PY,
           prop,
           colour = SO)) +
  geom_line(size = 2) +
  geom_text_repel(
    data = subset(cran_cites_all_areas_top_journals_by_year, 
                  PY == max(PY)),
    aes(label = str_wrap(SO_n, 40)),
    size = 4,
    nudge_x = 0,
    hjust = 0,
    segment.color = NA,
    direction = "y"
  ) +
   scale_x_continuous(breaks = min_y:max_y,
                     labels = c(min_y:2017, rep("", length(2018:max_y))),
                     limits = c(min_y, max_y)) +
  scale_y_continuous(labels = scales::percent) +
  xlab("Publication year") +
  ylab("Percentage of articles in that journal in that year") +
  theme_minimal() +
  theme(legend.position="none",
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

@@

# tally by year
archy_cran_cites_all_areas_top_journals_by_year <- 
  archy_cran_all_years %>% 
  filter(SO %in% top_journals_for_archy$SO) %>% 
  group_by(SO, PY) %>% 
  tally(sort = TRUE)  %>% 
  left_join(top_journals_for_archy_per_year,
            by = c('SO', 'PY')) %>% 
   mutate(prop = n.x / n.y) %>% 
  ungroup()

@@

min_y <- 2007
max_y <- 2040

all_archaeology_citing_R_over_time <- 
ggplot(archy_cran_cites_all_areas_top_journals_by_year,
       aes(PY,
           prop,
           colour = SO)) +
  geom_line(size = 2) +
  geom_point() +
  geom_text_repel(
    data = archy_cran_cites_all_areas_top_journals_by_year %>% 
      group_by(SO) %>% 
      slice(which.max(PY)) , 
    aes(label = SO_n),
    direction = "y",
    size = 5,
    nudge_x = 1,
    hjust = 0) +
  theme_minimal() +
  theme(legend.position="none",
        axis.text.x = element_text(angle = 90, 
                                   hjust = 1, 
                                   vjust = 0.5)) +
  scale_x_continuous(breaks = min_y:max_y,
                     labels = c(min_y:2017, 
                                rep("",
                                    length(2018:max_y))),
                     limits = c(min_y, max_y)) +
  scale_y_continuous(labels = scales::percent,
                     limits = c(0, 0.2)) +
  xlab("Publication year") +
  ylab("Percentage of articles in that journal")

@@

jas_cites_r <- 
archy_cran_cites_all_areas_top_journals_by_year %>% 
  filter(SO == "JOURNAL OF ARCHAEOLOGICAL SCIENCE") %>% 
  ggplot(aes(PY,
           prop)) +
  geom_point(size = 4) +
  geom_smooth(method = "lm",
              se = FALSE) +
  stat_poly_eq(aes(label =  paste(..eq.label.., 
                                  ..adj.rr.label..,
                                  sep = "~~~~")),
               formula = y~x, 
               parse = TRUE,
               size = 3,
               label.y.npc = 0.7) +
  stat_fit_glance(label.y.npc = 0.7,
                  size = 3,
                  method = "lm", 
                  method.args = list(formula = y ~ x),
                  geom = "text",
                  aes(label = paste("p-value: ",
                                    signif(..p.value.., 
                                           digits = 4)))) +
  xlab("year") +
  ylab("Proportion of articles citing R") +
  ggtitle(expression(paste("Articles citing R in ", 
                           italic("Journal of Archaeological Science")))) +
  theme_bw(base_size = 6)
```

@@

# combine main plot and subplot
all_archaeology_citing_R_over_time_and_subplot <- 
all_archaeology_citing_R_over_time  +
  annotation_custom(grob = ggplotGrob(jas_cites_r), 
                          xmin = 2030, xmax = 2040, 
                          ymin = 0.10, ymax = 0.17)
